---
title: "Data Exploratin and Deterministic Functions Assignment"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: cerulean
  
author: "Ruichen Wang"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r here packages install, echo = FALSE}
require(here)
```
```{r load the data, echo = FALSE}
dat_habitat = data.frame(read.csv(here("data","hab.sta.csv")))
```


### Terrain Histograms

```{r terrain histograms, echo = FALSE, fig.height = 10}
par(mfrow = c(3 , 1))
hist(dat_habitat$elev, breaks = 10,
     main = "Histogram of elevation",
     xlab = "elevation")
hist(dat_habitat$slope, breaks = 10,
     main = "Histogram of slope",
     xlab = "slope")
hist(dat_habitat$aspect, breaks = 15,
     main = "Histogarm of aspect",
     xlab = "aspect")
```

```{r visual estimation of linear model, echo = FALSE}

# Calculates the value of y for a linear function, given the coordinates
# of a known point (x1, y1) and the slope of the line.
line_point_slope = function(x, x1, y1, slope)
{
  get_y_intercept = 
    function(x1, y1, slope) 
      return(-(x1 * slope) + y1)
  
  linear = 
    function(x, yint, slope) 
      return(yint + x * slope)
  
  return(linear(x, get_y_intercept(x1, y1, slope), slope))
}

```

### Scatterplots

```{r scatterplots, echo = FALSE, fig.height = 10}
par(mfrow = c(3, 1))
plot(x = dat_habitat$elev, y = dat_habitat$ba.tot, 
     xlab = "elevation", ylab = "Total Basal Area",
     main = "Basal area and elevation", col = "blue", cex = 0.6)
curve(line_point_slope(x, x1 = 475, y1 = 20, slope = -0.01), add = TRUE, col = "green")

plot(x = dat_habitat$slope, y = dat_habitat$ba.tot,
     xlab = "slopt", ylab = "Total Basal Area",
     main = "Basal area and slope", col = "red", cex = 0.6)
curve(line_point_slope(x, x1 = 50, y1 = 30, slope = 0.05), add = TRUE, col = "green")

plot(x = dat_habitat$aspect, y = dat_habitat$ba.tot,
     xlab = "aspect", ylab = "Total Basal Area",
     main = "Basal area and aspect", col = "orange", cex = 0.6)
curve(line_point_slope(x, x1 = 175, y1 = 30, slope = 0.02), add = TRUE, col = "green")


```


### Assignment Questions
1. Figures see above

2. The distribution of elevation at the bird census sample sites is a shape of normal distribution. There are more sampling sites from 200-500 m, which is approximately in the middle of the range. Neither high- nor low-elevation has a high frequency. 

3. The slope is described as percent slope, ranging from 0 - 110, and is unitless. 

4. The distribution of the slope at the bird census sample sites is a shape of normal distribution. There are more sampling sites from 40-60% slope, which is approximately in the middle of the range. Neither too flat nor too steep of a slope has a high frequency. 

5. Aspect is the direction that a slope faces. The unit of aspect in this data set is degree, ranging from 0 to 360 degrees. 

6. The distribution of the aspect at the bird census sample sites is evenly distributed. The frequency of the aspects is about the same, and neither north-facing nor south-facing aspect is dominating. 

7. Figures see above

8. There is not a noticeable association. My linear model seems not fit the data very well because there is not a visually clear trend.
