---
title: "Probability and Frequentist Concepts"
author: "Ruichen Wang"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Probability of observing a count of exactly 3 successes in a binomial distribution with parameters n = 4 and p = 0.75
```{r}
dbinom(3,size = 4, prob = 0.75)
```

2. Probability of observing a count of 3 successes or fewer
```{r}
pbinom(3, size = 4, prob = 0.75)
```
or
```{r}
1 - dbinom(4, size = 4, prob = 0.75)
```

3. Probability of observing more than 3 successes in a binomial distribution with parameters n = 5 and p = 0.75
```{r}
1 - pbinom(3, size = 5, prob = 0.75)
```

4. Probability of observing a value of less than 1.2 from a normally-distributed population with mean = 2 and standard deviation = 2
```{r}
pnorm(1.2, mean = 2, sd = 2)
```

5. Probability of observing a value of greater than 1.2
```{r}
1 - pnorm(1.2, mean = 2, sd = 2)
```

6. Probability of observing a value between 1.2 and 3.2
```{r}
pnorm(3.2, mean = 2, sd = 2) - pnorm(1.2, mean = 2, sd = 2)
```

7. The shape of of histogram is approaching a skewed distribution to the left.

8. The shape of of histogram is approaching a normal distribution.

9. The shape of of histogram is approaching a normal distribution. The shape is narrower than the distribution in Q8.

10. Because for sample size of 1 the mean is the draw itself, whereas for sample size of 2 the mean lies between of the 2 draws, so there is a larger probability that the mean is normally distributed. 

11. The two main factors that determine the width of the sampling distribution of the mean are sample size and population standard deviation.

12. 
```{r}
25 * 25 * 25
```

13. 25B
